---
title: "Probabilistic_redesign_draft_site_selection"
author: "Keleigh Reynolds"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 2024 EPA Probabilistic Draw redesign
we are doing 80 sites over 2 years, with stratification by SO.



```{r begin-draw, message = FALSE, warning = FALSE}
library(spsurvey)
citation("spsurvey")

```


```{r begin-GRTS}
set.seed(22) #set seed for analysis

```

we are going to run it with unequal inclusion probability; stratified by eco-region Level III

# from previous draws

Survey Design: A Generalized Random Tessellation Stratified (GRTS) survey design for a linear resource was used.  The GRTS design includes reverse hierarchical ordering of the selected sites.

Multi-density categories:  Rivers are selected with unequal probability by stream order category: 1st, 2nd, 3rd, and 4th+ strahler orders.

Also used : 
"The coordinate reference system (CRS) for the sample frame should use an area-preserving projection such as Albers or UTM so that spatial distances are equivalent for all directions. Geographic CRS are not accepted." converted to UTM


```{r get-backround-files}

# example<-spsurvey::NE_Lakes
# prev_output<-read.csv("C:/Users/kareynol/New York State Office of Information Technology Services/SMAS - Site Selection/Probabilistic_epa_draw_data/Probabilistic/2018_2022Draw/2018_prob_draw.csv")

 # basin_path <- here::here("data/map_files") #for keleigh
 # basin <- rgdal::readOGR(
 #   dsn = basin_path,
 #   layer = "NHD_plus_2_basin",
 #   verbose = FALSE
 # ) rgdal is no longer supported by this version of R
 
 nhd<-sf::read_sf("data/map_files/NRSA2324.shp")

# # change coords to UTM for better resolution? as per spsurvey
# basin_shp <- sp::spTransform(
#   basin,
#   sp::CRS("+proj=utm +zone=18 +datum=WGS84"))


nhd_shp<-st_transform(nhd, crs = "+proj=utm +zone=18 +datum=WGS84") #works with the sf object



```



```{r}
#!fcode %in% c(#,#,# etc.)

#nhd_sf<-sf::st_as_sf(nhd_shp)
nhd_sf<-nhd_shp

nhd_sf<-nhd_sf %>% 
  dplyr::mutate(fcode=as.numeric(FCODE)) %>% 
  dplyr::filter(fcode != 33600|
                  fcode != 33601|
                  fcode != 33603|
                  fcode != 42000|
                  fcode != 42001|
                  fcode != 42002|
                  fcode != 42003|
                  fcode != 56600
                ) %>% 
  dplyr::filter(!dplyr::between(fcode,42800,42824)) %>% 
  dplyr::mutate(stream_order = as.numeric(STRAH_ORD)) %>% 
  dplyr::mutate(stream_order = dplyr::case_when(
    stream_order >= 4~4,
    TRUE ~ stream_order
  )) %>% 
  dplyr::filter(stream_order != 0) %>% 
  dplyr::filter(US_L3NAME != "")


nhd_sum<-as.data.frame(nhd_sf) %>% 
  group_by(US_L3NAME,stream_order) %>% 
  mutate(num= 1) %>% 
  summarise(sum = sum(num))

```



```{r using-epa-notes}
## set strata variable
nhd_sf<-st_as_sf(nhd_sf)
nhd_sf<-nhd_sf %>% 
  filter(stream_order>0) %>% 
  filter(!is.na(US_L3NAME))

stratum_var <- "US_L3NAME"
stratum_names <-unique(nhd_sf$US_L3NAME)

## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 150 sites statewide - so 16 per eco-region
caty_n2 <- lapply(1:9, function(x) c("1" = 5, "2" = 5, "3" = 3, "4" = 3))
names(caty_n2) <- unique(nhd_sf$US_L3NAME)
caty_n2

strata_n<-rep(16,9)
names(strata_n)<-unique(nhd_sf$US_L3NAME)
strata_n

### n2
samp_n150 <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n,
  #n_over = 0,
  stratum_var = "US_L3NAME",
  caty_n = caty_n2,
  caty_var = "stream_order",
  seltype = "unequal"
)

base150<-as.data.frame(samp_n150$sites_base)
write.csv(base150,"outputs/eco_region_redesign_150.csv")


```



```{r next-step-down-minimum - 60 sites total}
## set strata variable
stratum_var <- "US_L3NAME"
stratum_names <-unique(nhd_sf$US_L3NAME)

## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 60 sites statewide - so 6 per eco-region
caty_n <- lapply(1:9, function(x) c("1" = 2, "2" = 2, "3" = 1, "4" = 1))
names(caty_n) <- unique(nhd_sf$US_L3NAME)
caty_n

strata_n<-rep(6,9)
names(strata_n)<-unique(nhd_sf$US_L3NAME)
strata_n

### n2
samp_n50 <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n,
  #n_over = 0,
  stratum_var = "US_L3NAME",
  caty_n = caty_n,
  caty_var = "stream_order",
  #seltype = "unequal"
)

base50<-as.data.frame(samp_n50$sites_base)
write.csv(base50,"outputs/eco_region_redesign_50.csv")

```

```{r no strata}
## set strata variable
stratum_var <- "US_L3NAME"
stratum_names <-unique(nhd_sf$US_L3NAME)

## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 60 sites statewide - so 6 per eco-region
# caty_n <- lapply(1, function(x) c("1" = 2, "2" = 2, "3" = 1, "4" = 1))
# names(caty_n) <- unique(nhd_sf$US_L3NAME)
# caty_n
# 
# strata_n<-rep(6,9)
# names(strata_n)<-unique(nhd_sf$US_L3NAME)
# strata_n

### n2
samp_n150_no_strata <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = 150,
  #n_over = 0,
  #stratum_var = "US_L3NAME",
  #caty_n = caty_n,
  #caty_var = "stream_order",
  #seltype = "unequal"
)

base50<-as.data.frame(samp_n150_no_strata$sites_base)
write.csv(base50,"outputs/eco_region_redesign_150_no_strata.csv")

base50_sum<-base50 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))

nhd_sum_2<-as.data.frame(nhd_sf) %>% 
  group_by(stream_order) %>% 
  reframe(n=length(FRAME23_ID),
            km=sum(LENGTH_KM),
            miles = km / 0.62137119) %>% 
  ungroup() %>% 
  mutate(sum = sum(miles)) %>% 
  group_by(stream_order) %>% 
  mutate(proportion = miles/sum)

```

```{r no-strata-60-base}
### n2
samp_n60_no_strata <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = 60,
  #n_over = 0,
  #stratum_var = "US_L3NAME",
  #caty_n = caty_n,
  #caty_var = "stream_order",
  #seltype = "unequal"
)

base60_ns<-as.data.frame(samp_n60_no_strata$sites_base)
write.csv(base60_ns,"outputs/eco_region_redesign_60_no_strata.csv")

base60_sum<-base60_ns %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))

```

```{r stream-order-strata-no-eco}
## set strata variable
stratum_var <- "stream_order"
stratum_names <-unique(nhd_sf$stream_order)
strata_n_l <- c("1" = 45, "2" = 45, "3" = 30, "4" = 30)
names(strata_n_l) <- unique(nhd_sf[[stratum_var]])


## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 60 sites statewide - so 6 per eco-region
caty_n <- c("1" = 45, "2" = 45, "3" = 30, "4" = 30)
names(caty_n) <- unique(nhd_sf[[stratum_var]])


### n2
samp_n150_so_only <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n_l,
  #n_over = 0,
  stratum_var = "stream_order",
  #caty_n = caty_n,
  #caty_var = caty_n
)


base50<-as.data.frame(samp_n150_so_only$sites_base)
write.csv(base50,"outputs/eco_region_redesign_150_so_only.csv")

base50_sum<-base50 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))


```

```{r stream-order-proportion-strata-no-eco}
## set strata variable
stratum_var <- "stream_order"
stratum_names <-unique(nhd_sf$stream_order)
strata_n_l <- c("1" = 75, "2" = 33, "3" = 17, "4" = 19)
names(strata_n_l) <- unique(nhd_sf[[stratum_var]])

### n2
samp_n150_so_only_prop <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n_l,
  #n_over = 0,
  stratum_var = "stream_order",
  #caty_n = caty_n,
  #caty_var = caty_n
)


base50<-as.data.frame(samp_n150_so_only_prop$sites_base)
write.csv(base50,"outputs/eco_region_redesign_150_so_prop.csv")

base50_sum<-base50 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))


```




