---
title: "Probabilistic_redesign_draft_site_selection"
author: "Keleigh Reynolds"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 2024 EPA Probabilistic Draw redesign
we are going to do 80 sites stratified by stream order


```{r begin-draw, message = FALSE, warning = FALSE}
library(spsurvey)
citation("spsurvey")

```


```{r begin-GRTS}
set.seed(22) #set seed for analysis

```

we are going to run it with unequal inclusion probability; stratified by eco-region Level III

# from previous draws

Survey Design: A Generalized Random Tessellation Stratified (GRTS) survey design for a linear resource was used.  The GRTS design includes reverse hierarchical ordering of the selected sites.

Multi-density categories:  Rivers are selected with unequal probability by stream order category: 1st, 2nd, 3rd, and 4th+ strahler orders.

Also used : 
"The coordinate reference system (CRS) for the sample frame should use an area-preserving projection such as Albers or UTM so that spatial distances are equivalent for all directions. Geographic CRS are not accepted." converted to UTM


```{r get-backround-files}

 nhd<-sf::read_sf("data/map_files/NRSA2324.shp")
#this was taken from teh filegeo-database 
#L:\DOW\BWAM Share\data\streams\ArcGIS_Data\Hyrdo\NHD_Layers\EPA_NHD_PLUS_created_2023_03_01, and a shape file was created and loaded into this project

# # change coords to UTM for better resolution? as per spsurvey
# basin_shp <- sp::spTransform(
#   basin,
#   sp::CRS("+proj=utm +zone=18 +datum=WGS84"))


nhd_shp<-sf::st_transform(nhd, crs = "+proj=utm +zone=18 +datum=WGS84") #works with the sf object


```



```{r}
#!fcode %in% c(#,#,# etc.)

#nhd_sf<-sf::st_as_sf(nhd_shp)
nhd_sf<-nhd_shp

nhd_sf<-nhd_sf %>% 
  dplyr::mutate(fcode=as.numeric(FCODE)) %>% 
  dplyr::filter(FRAME23 =="Include")%>% 
  dplyr::mutate(stream_order = as.numeric(STRAH_ORD)) %>% 
  dplyr::mutate(stream_order = dplyr::case_when(
    stream_order >= 4~4,
    TRUE ~ stream_order)) %>% 
  dplyr::filter(stream_order != 0)

nhd_sf_less_4<-nhd_shp %>% 
  dplyr::mutate(fcode=as.numeric(FCODE)) %>% 
  dplyr::filter(FRAME23 =="Include")%>% 
  dplyr::mutate(stream_order = as.numeric(STRAH_ORD)) %>%
  dplyr::filter(stream_order != 0) %>% 
  dplyr::filter(stream_order <= 4)


nhd_sum<-as.data.frame(nhd_sf) %>% 
  group_by(stream_order) %>% 
  mutate(num= 1) %>% 
  summarise(sum = sum(num))



```


```{r stream-order-strata-no-eco-80-sites}
## set strata variable
stratum_var <- "stream_order"
stratum_names <-unique(nhd_sf$stream_order)
strata_n_l <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(strata_n_l) <- unique(nhd_sf[[stratum_var]])


## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 80 sites statewide
caty_n <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(caty_n) <- unique(nhd_sf[[stratum_var]])

n_over = c("1" = 96, "2" = 96, "3" = 64, "4" = 64)
names(n_over) <- unique(nhd_sf[[stratum_var]])

### n2
samp_n80 <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n_l,
  n_over =n_over,
  stratum_var = "stream_order",
  #caty_n = caty_n,
  #caty_var = caty_n
)


base80<-as.data.frame(samp_n80$sites_base)
overdraw<-as.data.frame(samp_n80$sites_over)

prob_24_25<-rbind(base80,overdraw)

write.csv(prob_24_25,"outputs/prob_draw_24_25.csv")

base80_sum<-base80 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))


```

```{r}
## set strata variable
stratum_var <- "stream_order"
stratum_names <-unique(nhd_sf$stream_order)
strata_n_l <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(strata_n_l) <- unique(nhd_sf[[stratum_var]])


## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 80 sites statewide
caty_n_2 <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(caty_n_2) <- unique(nhd_sf[[stratum_var]])

n_over = c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(n_over) <- unique(nhd_sf[[stratum_var]])

### n2
samp_n80_2 <- spsurvey::grts(
  sframe = nhd_sf,
  n_base = strata_n_l,
  n_over =n_over,
  stratum_var = "stream_order",
  #caty_n = caty_n,
  #caty_var = caty_n
)


base80<-as.data.frame(samp_n80_2$sites_base)
overdraw<-as.data.frame(samp_n80_2$sites_over)

prob_24_25_2<-rbind(base80,overdraw)

write.csv(prob_24_25_2,"outputs/prob_draw_24_25_smaller_overdraw.csv")

base80_sum_2<-base80 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))
```


```{r subset with long island legacy, smaller overdraw and no larger than 4th}
#legacy
options(scipen = 9999)
leg<-spsurvey::NE_Lakes_Legacy
#adding legacy site as per communication with Dan Rearick on 2/9/24


legacy_LI <- data.frame (stream_order  = c("3"),
                  latitude = c("4531702.45272212"),
                  longitude = c("691597.159403965"))
#weight of this one originally 42302.29222

legacy_LI<-sf::st_as_sf(legacy_LI, coords =c("longitude", "latitude"))
legacy_LI<-sf::st_set_crs(legacy_LI,"+proj=utm +zone=18 +datum=WGS84")


## set strata variable
stratum_var <- "stream_order"
stratum_names <-unique(nhd_sf$stream_order)
strata_n_l <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(strata_n_l) <- unique(nhd_sf[[stratum_var]])


## set caty variable
caty_var <- "stream_order"

## set caty sample sizes
#this is for 80 sites statewide
caty_n_2 <- c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(caty_n_2) <- unique(nhd_sf[[stratum_var]])

n_over = c("1" = 24, "2" = 24, "3" = 16, "4" = 16)
names(n_over) <- unique(nhd_sf[[stratum_var]])

### n2
samp_n80_3 <- spsurvey::grts(
  sframe = nhd_sf_less_4,
  n_base = strata_n_l,
  n_over =n_over,
  stratum_var = "stream_order",
  #caty_n = caty_n,
  #caty_var = caty_n
  legacy_sites = legacy_LI
)


# base80_3<-as.data.frame(samp_n80_3$sites_base)
# overdraw<-as.data.frame(samp_n80_3$sites_over)

combined <- sp_rbind(samp_n80_3)
write_sf(combined, here::here("outputs/legacy_24_25.shp"))


base80_sum_3<-base80_3 %>% 
  select(stream_order,siteID) %>%
  group_by(stream_order) %>% 
  summarise(n=(length(siteID)))
```
